---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rcdo

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/rcdo)](https://CRAN.R-project.org/package=rcdo)
<!-- badges: end -->

rcdo is a wrapper around [Climate Data Operators](https://code.mpimet.mpg.de/projects/cdo). 

## Installation

You can install rcdo from CRAN with 

```r
install.packages("rcdo")
```

or the development version of rcdo from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("eliocamp/rcdo")
```

Most operators are supported and are partially documented. 
The functions start with `cdo_` an the name of the operator (e.g. the selname operator is the `cdo_selname()` function)

## Example



```{r example}
library(rcdo)
cdo_use("packaged")  # use package version of cdo that can be installed with `cdo_install()`. 
ncep <- "hgt_ncep.nc"
```

The ymonmean operator computes monthly annual cycle. 
The rcdo function is `cdo_ymonmean()`

```{r}
ncep |> 
  cdo_ymonmean() 
```

The output just prints the command with a place holder output. 
Use `cdo_execute()` to actually run the command. 
If no outpuf file is specified, then the result is saved in a tempfile. 

```{r, message=FALSE}
ncep |> 
  cdo_ymonmean() |> 
  cdo_execute()
```

Operators can be chained. 
Lets select just the Southern Hemisphere first.

```{r}
ncep |> 
  cdo_sellonlatbox(lon1 = 0, lon2 = 360, lat1 = -90, lat2 = 0) |> 
  cdo_ymonmean() 
```

Now also select the 500 hPa level 

```{r}
ncep |> 
  cdo_sellonlatbox(lon1 = 0, lon2 = 360, lat1 = -90, lat2 = 0) |> 
  cdo_sellevel(level = 500) |> 
  cdo_ymonmean() 
```


```{r}
ncep |> 
  cdo_sellonlatbox(lon1 = 0, lon2 = 360, lat1 = -90, lat2 = 0) |> 
  cdo_sellevel(level = 500) |> 
  cdo_ymonmean() 
```

## Prior art

The [ClimateOperators](https://github.com/markpayneatwork/ClimateOperators) package also wrapps CDO, but it's approach is different. 
Instead of wrapping each operator as its own function with parameters as arguments, it provides a generic `cdo()` function that runs the operators that the user needs to write as strings. 
Instead of 

```{r, eval=FALSE}
ncep |> 
  rcdo::cdo_sellonlatbox(lon1 = 0, lon2 = 360, lat1 = -90, lat2 = 0) 
```

one would write 
 
```{r, eval=FALSE}
ClimateOperators::cdo("sellonlatbox,0,360,-90,0", ncep, output_file)
```
